Useful R tidbits: Social Networks

### 2022-03-26 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
General updates

```r
igraph::degree(trade2007.ig, v = V(trade2007.ig)[c("China", "Canada", "United Kingdom", "Denmark")])
         China         Canada United Kingdom        Denmark 
           364            364            364            362 
```

@@

@@

# Statnet-only: Gould-Fernandez Brokerage
temp <- data.frame(brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)$z.nli)
climinfl.nodes <- climinfl.nodes %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)

@@

mat_climinfl <- as.matrix(as_adjacency_matrix(climinfl.ig))  # not " attr='weight'"
mat_climinfl_sq <- t(mat_climinfl) %*% mat_climinfl
climinfl.nodes$rc <- diag(mat_climinfl_sq) / rowSums(mat_climinfl_sq)
# replace missing values with 0
climinfl.nodes$rc <- ifelse(is.nan(climinfl.nodes$rc), 0, climinfl.nodes$rc)
climinfl.nodes$dc <- 1 - climinfl.nodes$rc

@@

Requires a directed network with a vertex attribute used for grouping. Returns a structure with a lot of information is returned; tutorial refers mainly to `znli` containing the following roles:

@@

>>>>

```r
> igraph::degree(trade2007.ig, v = V(trade2007.ig)[c("China", "Canada", "United Kingdom", "Denmark")])
         China         Canada United Kingdom        Denmark 
           364            364            364            362 
```

This can also be done by index, as in:

```r
> igraph::degree(flo_ig, v = 1:3)
Acciaiuoli    Albizzi  Barbadori 
         2          6          4 
```

Statnet degrees aren't named:

```r
> sna::degree(flo_stat)
 [1]  2  6  4  6  6  2  8  2 12  2  6  0  6  4  8  6
> which(sna::degree(flo_stat) == 0)
[1] 12
```

@@

*Note*: a graph is fully connected if its number of components is 1. igraph returns this as the `no` parameter of `igraph::components()`; it appears that statnet has no parallel function, but if the number of isolates is 0, the graph is connected (`sna::isolates()`).

@@

# Statnet-only: Gould-Fernandez Brokerage
# replace ATTR with the vector of the desired attribute, such as
# `climinfl.nodes$orgtype5`
temp <- data.frame(brokerage(climinfl.stat, cl = ATTR)$z.nli)
climinfl.nodes <- climinfl.nodes %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)

@@

mat_climinfl <- as.matrix(as_adjacency_matrix(climinfl.ig))  # not " attr='weight'"
mat_climinfl_sq <- t(mat_climinfl) %*% mat_climinfl
# alternately:
mat_climinfl <- as.matrix.network(climinfl.stat, attr = "weight")
diag(mat_climinfl) <- 0
mat_climinfl_sq <- mat_climinfl %*% mat_climinfl

climinfl.nodes$rc <- diag(mat_climinfl_sq) / rowSums(mat_climinfl_sq)
# replace missing values with 0
climinfl.nodes$rc <- ifelse(is.nan(climinfl.nodes$rc), 0, climinfl.nodes$rc)
climinfl.nodes$dc <- 1 - climinfl.nodes$rc

@@

Requires a directed network with a vertex attribute used for grouping. Returns a structure with a lot of information; tutorial refers mainly to `znli` containing the following roles:

@@

## Week 6

(All statnet unless otherwise specified)

Calculating structural equivalence ("SE clusters"): look for nodes that have the same pattern of ties with the same neighbors, like siblings of a parent. Unexplained distance functions include "hamming", "correlation", "gamma". The `sedist()` function ignores edge values (weights?).

```r
flomarr.se <- equiv.clust(flomarr.stat, equiv.fun = "sedist", method = "hamming", mode = "graph")
```

The output of `equiv.clust()` can be plotted as a "Cluster Dendogram":

```r
flomarr.se <- equiv.clust(flomarriage, equiv.fun = "sedist", method = "hamming", mode = "graph")
plot(flomarr.se,labels = flomarr.se$glabels)
```

Annoyingly, the different `cluster.method`s "single, average, or ward.D" are not explained in the tutorial. The default is "complete".

`rect.hclust()` (visually) "cuts" the diagram at a given height, making separate clusters:

```r
plot(flomarr.se, labels = flomarr.se$glabels)
rect.hclust(flomarr.se$cluster, h = 9)
```

```r
flomarr.se <- equiv.clust(flomarriage, equiv.fun = "sedist", cluster.method = "single",
                          method = "hamming", mode = "graph")
plot(flomarr.se,labels = flomarr.se$glabels)
